---
title: "ggplot"
author: "Andrew Witherspoon"
date: "9/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

dir <- "/Users/andrew/Course4Week2"
setwd(dir)
```

For these examples, we will be using the mpg dataset:
```{r}
library(ggplot2)
str(mpg)
```

This data frame uses clearly labled factors to subset the data, which is important for the ggplot style of graphing.


A simple plot, using "quick plot"" - qplot()
```{r}
qplot(displ, hwy, data = mpg)
```

Now let's do it again, but use the color call to subset by shape, on the drv variable:
```{r}
qplot(displ, hwy, data = mpg, shape = drv)
```

You can specify the way the data is presented using the geom argument.  qplot will attempt to guess what geom you want depending on the input:

# both x and y supplied = scatterplot
qplot(mpg, wt, data = mtcars)
# just x supplied = histogram
qplot(mpg, data = mtcars)
# just y supplied = scatterplot, with x = seq_along(y)
qplot(y = mpg, data = mtcars)

Or, of course, you can specify.  Here is geom = "point":
```{r}
qplot(displ, hwy, data = mpg, geom = "point")
```

And geom = "smooth":
```{r}
qplot(displ, hwy, data = mpg, geom = "smooth")
```

Or - we could do both using geoom = c("point", "smooth"):
```{r}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```

As stated above, supplying only one variable will default to the histogram geom:
```{r}
qplot(hwy, data = mpg, fill = drv)
```

Instead of using different shapes or colors to subset data by a factor variable, we can use facets (similar to a lattice plot):
```{r}
qplot(displ, hwy, data = mpg, facets = .~drv)
```

Here, we do facets with a histogram:
```{r}
qplot(hwy, data= mpg, facets = drv~., binwidth = 2)
```

The format for the facets argument is a variable to the left of the ~ to indicate rows, and a variable on the right of the ~ to indicate columns, to create a matrix of plots.

In each of these cases, there is only one row/column, so a dot is used on one side of the ~

Here is one with two variables for the facets argument:
```{r}
qplot(displ, hwy, data = mpg, facets = cyl~drv)
```
